Nanophotonics with optical phonons in two dimensions

### 2018-11-11 paper.tex
text basically done. need to add figures and an equation now.

In this Letter, we develop core results in the nascent field of optical phononics in two-dimensional materials by deriving a universal form for the dielectric function of a polar monolayer, incorporating spatially nonlocal effects, which is specified in fully in terms of the dispersion relation of LO and TO phonons in two dimensions. Using these parameters for important monolayers such as hexagonal boron nitride, we present the confinement and propagation losses of these modes, identifying the frequency region where they should be most easily detected. We... [insert nicest results here].

@@

where 'L' denotes longitudinal polarization, $|0_{\mathbf{q}L}\rangle$ ($|1_{\mathbf{q}L}\rangle$) denotes a state with no (one) longitudinal phonon of wavevector $\mathbf{q}$. The frequency $\omega_{\mathbf{q}L}$ in the denominator, as in the case of bulk phonons, is the frequency of the longitudinal phonon of wavevector $\mathbf{q}$ prior to considering LO-TO splitting [Needs reference!]. This is the consistent with the fact that LO-TO splitting is a collective effect arising from Coulomb interactions and the fact that the equation above represents a single-particle susceptibility. Coulomb interactions are accounted for in the random phase approximation, and to include them in the single-particle response would be double-counting. The component $qq$ in the response tensor denotes a pair of directions parallel to the wavevector.  

@@

We now move to analyze t

\subsection{Spontaneous emission}

\subsection{Spontaneous and stimulated electron energy loss}

>>>>

In this Letter, we develop core results in the nascent field of optical phononics in two-dimensional materials by deriving a universal form for the dielectric function of a polar monolayer, incorporating spatially nonlocal effects, which is specified in fully in terms of the dispersion relation of LO and TO phonons in two dimensions. Using these parameters for important monolayers such as hexagonal boron nitride, we present the confinement and propagation losses of these modes, identifying the frequency region where they should be most easily detected. We then show that these modes essentially provide the ultimate limit of light-matter interaction between emitters and polar materials, showing that for atom-like emitters, their spontaneous decay can be enhanced seven orders of magnitude by the emitter-LO phonon coupling. For free-electrons, we show a substantial probability of multiple LO phonon excitation can occur.

@@

where 'L' denotes longitudinal polarization, $|0_{\mathbf{q}L}\rangle$ ($|1_{\mathbf{q}L}\rangle$) denotes a state with no (one) longitudinal phonon of wavevector $\mathbf{q}$. The frequency $\omega_{\mathbf{q}L}$ in the denominator, as in the case of bulk phonons, is the frequency of the longitudinal phonon of wavevector $\mathbf{q}$ prior to considering LO-TO splitting [Needs reference!]. This is the consistent with the fact that LO-TO splitting is a collective effect arising from Coulomb interactions and the fact that the equation above represents a single-particle susceptibility. Coulomb interactions are accounted for in the random phase approximation, and to include them in the single-particle response is an uncontrolled double-counting. The component $qq$ in the response tensor denotes a pair of directions parallel to the wavevector.  

@@

The extreme confinement of electromagnetic fields offered by the 2D LO phonon presents an opportunity for quantum optical applications in which one seeks to couple an external emitter such as an atom, molecule, defect, or artificial atomic system to electromagnetic fields. Applications of these couplings can be to realize ultra-bright single- or two-photon sources, or realize strong-coupling and the associated phenomenology of Rabi oscillations and polaritons, or resolve spectroscopically "forbidden" transitions. In Figure 3, we show key figures of emitter-phonon coupling in polar monolayers.

In Fig. 3a, we show the enhancement (or Purcell factor) $F_p$ of spontaneous emission of 2D LO phonons at frequency $\omega_0$ by an external emitter at different distances $d$ away from the surface. The rate of spontaneous emission $\Gamma$ into 2D LO phonons obeys
\begin{equation}
\Gamma = \frac{3c^3}{2\omega_0^3}\int dq~q^2e^{-2qz}\text{Im }r_p(q,\omega_0) \times \Gamma_0 \equiv F_p(\omega_0)\Gamma_0,
\end{equation}
where $\Gamma_0$ is the free-space spontaneous emission rate. As can be seen from Fig. 3a, for an emitter position 5 nm away from the surface, the Purcell factor can be very high, corresponding to over a million-fold increase in the rate of 2D LO phonon emission compared to photon emission into free-space. While this rate-enhancement is already quite high, it is a natural question what the highest possible spontaneous-emission rate enhancement is as a function of frequency. Recently, material-geometry-independent limits for spontaneous emission enhancement have been derived. In particular, the maximum rate of spontaneous emission enhancement is bounded by
\begin{equation}
...
\end{equation}
This is plotted in Fig. 3b for the case of monolayer hBN.

Finally, we address one manner of externally probing 2D LO phonons without the need for SNOM or emitters: with electron energy losses in free electrons. As the energy losses under consideration will be on the order of 200 meV for single phonon excitation in monolayer hBN, the electron energy spread which determines the width of EELS peak (including the zero-loss peak) must be substantially smaller than 200 meV. Fortunately, this is possible in high-resolution EELS, which is particularly useful for probing surface phenomena, and has been applied to measure the dispersion relation of plasmons in 2D or quasi-2D systems. Through advances in monochromation, electrons have also been recently used to probe optical phonons in bulk. In Fig. 4, we show the EELS probability for electrons with energies on the order of 100 keV traversing a polar monolayer . Because of the strong coupling between matter and LO phonons shown in Fig. 3, one expects multiple LO phonon excitation to be possible. Thus, this effect must be taken into account in the EELS. 
\begin{equation}
...
\end{equation}

The ability to probe low-loss and highly confined electromagnetic modes associated with 2D LO phonons in polar materials provides a new platform for nanophotonics in the mid- and far-infrared spectral range. The identification of the phonon polariton of bulk and thin-film geometries with the 2D LO phonon made in this manuscript may one to allow extend the rich phenomenology of optical phonons to nanophotonic applications. It also may lead to useful new ways to study LO phonons, arising from the fact that 2D LO phonons, unlike their 3D counterparts, have their electromagnetic energy extend a considerable distance from the material boundary. Due to the strong electromagnetic interactions between emitters and 2D LO phonons shown in this manuscript, it may be possible to design interesting new hybrid states of matter and phonons based on quantum electrodynamical strong coupling. The highly confined optical phonon waves in polar monolayers may also provide interesting new opportunities in near-field radiative heat transfer, in which it has been long known that thin-film surface phonon polaritons play a critical role. An important avenue of future study would be the \textit{ab initio} calculation of lifetimes of 2D LO phonons associated with three-phonon processes and electron-phonon interactions. It would be of great interest to study the prospects of techniques such as isotopic purification and cryogenic temperatures for reducing the decay of these 2D LO phonons.
